---
title: "Final Project"
subtitle: "Data Science for Biologists, Spring 2021"
author: "Vinny Varghese"
output:
  html_document: 
    theme: sandstone 
    highlight: breezedark 
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# DON'T CHANGE THE NEXT LINE, except to add NEW options if you have any.
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	include = TRUE
)
# Load all libraries you use HERE:  
library(tidyverse)
library(broom)
library(modelr)


```

<br><br>

## Introduction

**The data set that I chose was from kaggle.com, and it is called heart.csv. This dataset analyzes heart attacks and some of the factors that goes into predicting if a person will have a heart attack. Since I want to go into the medical field, this data set caught my eye. If using the techniques we learned in class this year can do an accurate job at predicting heart attacks, it will bode well for the future of the medical field.**

This is the link to the [Heart Attack Dataset](https://www.kaggle.com/rashikrahmanpritom/heart-attack-analysis-prediction-dataset)

```{r read-examine-data}
heart <- read_csv("varghese_heart.csv") 
heart

dplyr::glimpse(heart)   

```

<br><br>

## Question 1

#### According to healthline.com, "As you age, the vascular system changes. This includes your heart and blood vessels. In the blood vessels, there's a reduction in elastic tissue in your arteries, causing them to become stiffer and less compliant. As a result your blood pressure increases." Knowing this information, can this same knowledge be applied for the variable of resting blood pressure in this dataset?(Does resting blood pressure increase as you get older?)

### Methods

```{r Question 1 code}
heart %>% 
  select(age, trtbps) %>% #use select to keep only age and resting blood pressure to find the relationship  

ggplot() + 
  aes(x = age, y = trtbps) + #make age the x variable and trtbps the y variable  
  geom_point(size = 2, 
             color = "springgreen2") + #geom_point should be used for the wide variety of points in the data and make it a size that can be visibly seen  
  geom_smooth(method = lm) + #use geom_smooth to have a line of best fit to show the overall strength of the relationship between the two variables  
  labs(x = "Age", y = "Resting Blood Pressure(mmHg)")
  

```

### Answer

As seen by the scatter plot visual, there is a fairly strong relationship between these variables. Therefore, we can conclude that resting blood pressure tends to increase with age.

<br><br>

## Question 2

#### Does this dataset show a decrease in maximum heart rate as age increases?

### Methods

```{r Max HR vs Age}
heart %>% 
  select(age, thalachh) %>% #make sure to keep only the age and the maximum heart rate(thalachh) 
  
  ggplot(aes(x = age, y = thalachh)) + #age is the independent variable(x) and we are trying to see how max heart rate changes over time(y) 
  geom_point(size = 2, 
             color = "darkorchid1") + 
  geom_smooth(method = lm) + #use geom_smooth to get a line of best fit 
  labs(x = "Age", y = "Maximum Heart Rate(BPS)")


```

### Answer

This dataset does indeed show that there is a negative relationship between age and maximum heart rate. This confirms the general belief of the relationship of maximum heart rate of each individual over time.

<br><br>

## Question 3

#### Using Linear regression, to what extent does age explain variation in cholesterol? Is there a significant relationship between these variables(use scatterplot)?

### Methods

```{r perform model_selection and visualize}
fitted_model <- lm(chol ~ age, data = heart) #build a linear model with lm and define it

summary(fitted_model) #examine the model output

broom::tidy(fitted_model) #use this to get the main values to evaluate 

modelr::rsquare(fitted_model, heart) #find rsquared value 

heart %>% 
  select(chol, age) %>% 
  
  ggplot(aes(x = age, y = chol)) + 
  geom_point(size = 2, 
             color = "darkorange") + 
  geom_smooth(method = lm) +
  labs(x = "Age", y = "Cholesterol(mg/dl)")


```

### Answer

Although there is a relationship between between cholesterol level and age, it is fairly weak. The model that we built explains that 4.6% of variation in cholesterol is explained by age which isn't significant.

<br><br>

## Question 4

#### Which sex has the highest average cholesterol level according to this dataset?

### Methods

```{Cholesterol and Age code}      
heart %>% 
  group_by(sex) %>% #group variables by sex 
  summarize(mean_chol = mean(chol)) %>% #use the summary function to make a new column and get rid of any columns not used in calculation 
  
  ggplot(aes(x = sex, y = mean_chol, fill = sex)) + 
  geom_col(alpha = 0.7, 
           fill = c("deeppink", "dodgerblue1"), 
           width = 0.2) + #use geom_col to have two distinct variables and use an alpha value that makes each bar visible. Pink color = female, Blue = male 
  labs(x = "Sex", y = "Mean Cholesterol(mg/dl)", title = "Average Cholesterol between Genders")

```


### Answer

By comparing these two genders, we can see that women have a higher average cholesterol than men.

### Have a great summer :)
